---
output:
  html_document: default
  pdf_document: default
---

---
title: "Test of whether kinetic constants can be obtained with BioMark"
output: html_notebook
---

Copyright (C) 2018  Martin Engqvist Lab




## Experiment
.....

Experimental data collected on 181218 (DDMMYY) in Petter Wolls group at Karolinska using a BioMark instrument.

Assay conditions, oxidases:


20mM HEPES, pH 7.4

50uM Amplex Red

0,1U/ml HRP

0,1uM fluorescein

0,01mg/ml BSA

0,01% Triton-X-100 and Tween20


```{r setup}

#make sure the packages are installed
wants <- c('dplyr', 'reshape2', 'ggplot2', 'tidyverse', 'assertthat')
has   <- wants %in% rownames(installed.packages())
if(any(!has)) install.packages(wants[!has], repos="http://ftp.acc.umu.se/mirror/CRAN/")

#load all the packages
lapply(wants, require, character.only = TRUE)


#find the "code" directory, this should be one folder away from the project root
while (basename(getwd()) != 'code'){
  setwd("../")
  if (getwd() == '/'){
    quit(status = 1)
  }
}

#go one folder up and set as project root directory
setwd("../")
PROJ <- getwd()
setwd(paste0(PROJ, '/data/'))

#set folder paths relative to the project root
CURRENT <- getwd()
DATA <- getwd()
RAW_EXTERNAL <- paste0(DATA, '/raw_external/')
RAW_INTERNAL <- paste0(DATA, '/raw_internal/')
INTERMEDIATE = paste0(DATA, '/intermediate/')
FINAL = paste0(DATA, '/final/')

RESULTS <- paste0(PROJ, '/results/')
FIGURES <- paste0(RESULTS, 'figures/')
PICTURES <- paste0(RESULTS, 'pictures/')

```


## Define variables specific for this plate and array
```{r}
THIS_PROJECT_FOLDER <- 'flex_six_km_test'
PLATE_NUMBER <- 2
ARRAY_NUMBER <- 1
DATA_FOLDER <- sprintf('Plate_%s_Array_%s', PLATE_NUMBER, ARRAY_NUMBER)
FAM_DATA <- '1530024054_plate_2_array_1_fam.tsv'
ROX_DATA <- '1530024054_plate_2_array_1_rox.tsv'

MINUTES_PER_CYCLE <- 0.5 # how many minutes elapse between each cycle

R2_CUTOFF <- 0.8
CORR_CUTOFF <- 0.8
TOLERANCE_AT_ZERO <- 0.05
TOLERANCE_AT_VMAX <- 0.05

SLOPE_FIRST_TIMEPOINT <- 4
SLOPE_LAST_TIMEPOINT <- 20


# load helper functions
source(paste0(PROJ, '/code/', 'flex_six_helper_functions.R'))
print(paste0('helper script version: ', VERSION))
```


Setup plotting preferences

```{r}
font_size = 6
plt_height = 65
plt_width = 89
line_size = 0.5
line_width <- 0.5
error_size = 0.25
pd = position_dodge(0.9)

gen_col = '#141A46'
nat_col = '#EC8B5E'
neg_col = '#868686'



ggplot_theme <- theme(text = element_text( size=font_size), 
                      axis.text.x = element_text(angle = 0, hjust = 1, size=font_size),
                      axis.text.y = element_text(size=font_size),
                      axis.line = element_line(color = "black", size=line_size),
                      axis.ticks.x = element_line(color = 'black', size=line_size),
                      axis.ticks.y = element_line(color = 'black', size=line_size),
                      #legend.text = element_text( size=font_size),
                      legend.position = c(0, 1),
                      legend.justification = c(0, 1),
                      legend.background = element_blank(),
                      legend.direction = 'horizontal',
                      legend.text=element_text(size=font_size),
                      plot.title = element_text( size=font_size),
                      panel.background = element_blank(),
                      panel.border = element_rect(color = 'black', size=line_size),
                      #panel.margin = 
                      panel.grid.major = element_blank(), 
                      panel.grid.minor = element_blank(),
                      panel.grid.major.x = element_blank(), 
                      panel.grid.major.y = element_blank(), 
                      panel.grid.minor.x = element_blank(), 
                      panel.grid.minor.y = element_blank() 
)
```

## Setup

Let's input the experimental paramenters for sample and assay sides

```{r}
assay.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, sprintf('plate%s_array%s_assay.tsv', PLATE_NUMBER, ARRAY_NUMBER), sep='/'), sep='\t', header=TRUE, na.strings="")

sample.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, sprintf('plate%s_array%s_sample.tsv', PLATE_NUMBER, ARRAY_NUMBER), sep='/'), sep='\t', header=TRUE, na.strings="")


print(assay.df)
print(sample.df)
```




"Mix" the wells from input assay and sample data to do in silico what happens in the chip

```{r}
mixed.df <- mix_wells(assay.df, sample.df)

head(mixed.df)
tail(mixed.df)
```




## Load data


```{r}
# read the experimental data obtained from the BioMark
rox.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, ROX_DATA, sep='/'), sep='\t', header=TRUE, na.strings="")
fam.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, FAM_DATA, sep='/'), sep='\t', header=TRUE, na.strings="")

# combine data 
all.df <- combine_data(mixed.df, rox.df, fam.df)

# Add in the time at which each read took place
all.df$minutes <- all.df$cycle * MINUTES_PER_CYCLE

head(all.df)
```


Take a look at the frame and remove NA if required

```{r}
summary(all.df)

# replace all na values with zero to allow for analysis despite missing values
all.df[is.na(all.df)] <- 0

```


Let's export this to a flatfile for future use.

```{r}
write.table(all.df, file=paste(FINAL, THIS_PROJECT_FOLDER, sprintf('%s_flex_six_plate_%s_array_%s_raw_data.tsv', gsub(' ', '_', unique(all.df$assay_enzyme)), PLATE_NUMBER, ARRAY_NUMBER), sep='/'), sep='\t', row.names=FALSE)
```



## Plotting results
Time to plot!


Plot rox and fam values of first and last cycle

```{r}

plot_raw_data_heatmaps(all.df, min=0, max=65000)

```




First we plot the raw kinetics observed in all the wells

```{r}
ggplot(all.df, aes(x=minutes, y=rox_value, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Raw data measured over time, plate %s, array %s', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab('Time (minutes)') + 
    ylab('Fluorescence') +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```




Convert fluorescence to concentration of product using standard curve

```{r}
concentration.df <- fluorescence_to_concentration(all.df, chip_type='flex_six')

head(concentration.df)

in_data <- concentration.df
ggplot(in_data, aes(x=minutes, y=resorufin_value, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Raw data measured over time, plate %s, array %s', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab('Time (minutes)') + 
    ylab(sprintf('Resorufin concentration (%s)', unique(in_data$resorufin_value_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```




Make a subset which should be the linear range, plot again

```{r}
all.df.slope.subset <- concentration.df[concentration.df$minutes >= SLOPE_FIRST_TIMEPOINT & concentration.df$minutes <= SLOPE_LAST_TIMEPOINT, ]

in_data <- all.df.slope.subset
ggplot(in_data, aes(x=minutes, y=resorufin_value, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, plate %s, array %s', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab('Time (minutes)') + 
    ylab(sprintf('Resorufin concentration (%s)', unique(in_data$resorufin_value_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```



Subtract fluorescence from the no enzyme control

```{r}
# remove fluorescence from the no enzyme control
all.df.slope.subset.combined <- subtract_no_enz_fluorescence(all.df.slope.subset)

head(all.df.slope.subset.combined)

in_data <- all.df.slope.subset.combined
ggplot(in_data, aes(x=minutes, y=resorufin_value_norm, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, plate %s, array %s, normalized data', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab('Time (minutes)') + 
    ylab(sprintf('Resorufin concentration (%s)', unique(in_data$resorufin_value_norm_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```




Fit a linear regression for all data between the first and last points to get the slope.

```{r}
# Fit a linear regression for all data between the first and last points to get the slope.
slopes.df <- compute_slopes(all.df.slope.subset.combined)

head(slopes.df)
tail(slopes.df)
```




A look at the slope data as a heatmap and assess the fit of the linear model.

```{r}

### Plotting slope data and the fit of the model

plot_slope_data_heatmaps(slopes.df)


```


Remove all enzyme concentrations that have poor R2 values (below a set cutoff)

```{r}

bundled.slopes.nonegs.df <- filter_r2(slopes.df, r2_cutoff=R2_CUTOFF, corr_cutoff=CORR_CUTOFF)


head(bundled.slopes.nonegs.df)
```




Calculate spearman correlation between substrate concentration and the measured slope (the kinetic slope)

```{r}

bundled.slopes.nonegs.spearman.df <- calculate_slope_and_substrate_correlation(bundled.slopes.nonegs.df)

# look at the data
head(bundled.slopes.nonegs.spearman.df)

# plot it
barplot(bundled.slopes.nonegs.spearman.df$slope_spearman_rho, 
        names=interaction(bundled.slopes.nonegs.spearman.df$final_enzyme_conc, c('repeat'), bundled.slopes.nonegs.spearman.df$repeat_num),
        ylim=c(-1, 1),
        xlab='final_enzyme_conc',
        ylab='spearman rho',
        main='Which enzyme conc. is best?\n(corr. of substrate conc. and kinetic slope)',
        las=3)

```




For each repeat choose the enzyme concentration with maximal spearman rho

```{r}

bundled.slopes.nonegs.spearman.max.df <- select_best_samples(bundled.slopes.nonegs.spearman.df)

# look at the data
head(bundled.slopes.nonegs.spearman.max.df)

# plot it
barplot(bundled.slopes.nonegs.spearman.max.df$slope_spearman_rho, 
        names=interaction(bundled.slopes.nonegs.spearman.max.df$final_enzyme_conc, c('repeat'), bundled.slopes.nonegs.spearman.max.df$repeat_num),
        ylim=c(-1, 1),
        xlab='final_enzyme_conc',
        ylab='spearman rho',
        main='Selected enzyme concentrations',
        las=3)
```




Unbundle the dataset (turn the lists into column values again)

```{r}

bundled.slopes.nonegs.spearman.max.filtered.unbundled.df <- unbundle_data(bundled.slopes.nonegs.spearman.max.df)

head(bundled.slopes.nonegs.spearman.max.filtered.unbundled.df)
```




Subtract the zero substrate slope from all others

```{r}
# Subtract the zero substrate slope from all others
final.data.df <- subtract_no_substrate_slope(bundled.slopes.nonegs.spearman.max.filtered.unbundled.df)

# look at the data 
head(final.data.df)

# plot it
in_data <- final.data.df
ggplot(in_data, aes(x=final_substrate_conc, y=activity))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Activity, plate %s, array %s, background-subtracted data', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate conentration (%s)', unique(in_data$final_substrate_unit))) + 
    ylab(sprintf('Activity (%s)', unique(in_data$activity_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))


```



Need to convert from uM per minute to micro moles per minute

```{r}
final.data.moles.df <-concentration_to_moles(final.data.df, chip_type='flex_six')

head(final.data.moles.df)
```



Normalize to per mg protein
```{r}
final.data.moles.protein.df <- normalize_to_protein(final.data.moles.df, chip_type='flex_six')

head(final.data.moles.protein.df)
```


Make a plot of the new values
```{r}
# plot it
in_data <- final.data.moles.protein.df
ggplot(in_data, aes(x=final_substrate_conc, y=activity_moles_protein))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Activity, plate %s, array %s', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate conentration (%s)', unique(sample.df$sample_substrate_unit))) + 
    ylab(sprintf('Activity (%s)', unique(in_data$activity_moles_protein_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))


```




Trim off bad values at the very low and very high substrate concentrations
```{r}

# Remove bad values (low concentrations where no activity is seen, and high values where activity drops below max)
final.data.filtered.both.df <- trim_data_range(final.data.moles.protein.df, tolerance_at_zero=TOLERANCE_AT_ZERO, tolerance_at_max=TOLERANCE_AT_VMAX)

# look at the data
head(final.data.filtered.both.df)


# plot it
in_data <- final.data.filtered.both.df
ggplot(in_data, aes(x=final_substrate_conc, y=activity_moles_protein))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, plate %s, array %s, filtered data', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate conentration (%s)', unique(sample.df$sample_substrate_unit))) + 
    ylab(sprintf('Activity (%s)', unique(in_data$activity_moles_protein_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))



```



Get rid of extreme outliers

```{r}
# remove extreme outliers
final.data.filtered.both.no_outliers.df <- remove_outliers(final.data.filtered.both.df)

# look at the data
head(final.data.filtered.both.no_outliers.df)


# plot it
ggplot(final.data.filtered.both.no_outliers.df, aes(x=final_substrate_conc, y=activity_moles_protein))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, plate %s, array %s, outliers removed', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate conentration (%s)', unique(in_data$final_substrate_unit))) + 
    ylab(sprintf('Activity (%s)', unique(in_data$activity_moles_protein_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```



Export for future use

```{r}
write.table(final.data.filtered.both.no_outliers.df, file=paste(FINAL, THIS_PROJECT_FOLDER, sprintf('%s_flex_six_plate_%s_array_%s_final_data.tsv', gsub(' ', '_', unique(all.df$assay_enzyme)), PLATE_NUMBER, ARRAY_NUMBER), sep='/'), sep='\t', row.names=FALSE)
```




Plot slope vs substrate concentration, estimate Km and Vmax by fitting to the Michaelis-Menten equation

```{r}

kinetic_data <- calculate_kinetics(final.data.filtered.both.no_outliers.df)
mm_fit.df <- data.frame(xFit=kinetic_data$xFit, yFit=kinetic_data$yFit)
km <- kinetic_data[[3]]
vmax <- kinetic_data[[4]]


in_data <- final.data.filtered.both.no_outliers.df

p1 <- ggplot(in_data, aes(x=final_substrate_conc, y=activity_moles_protein))+
    theme_bw() +
    ggplot_theme +
    geom_point(cex=1.5, fill='red', color='red') + 
    geom_line(data=mm_fit.df, aes(x=xFit, y=yFit)) + 
    ggtitle(sprintf('Km plot, plate %s, array %s', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('%s (%s)', unique(in_data$sample_substrate), unique(in_data$final_substrate_unit))) + 
    ylab( expression(paste('Activity (', mu, 'mol mg'^{-1}, 'min'^{-1}, ')') ) ) +

    annotate("text", 
            x = min(in_data$final_substrate_conc) + max(in_data$final_substrate_conc)*0.2, 
            y = min(in_data$activity_moles_protein) + max(in_data$activity_moles_protein)*0.05, 
            label = as.expression(bquote(italic(V)[max] ~ '=' ~ .(formatC(vmax, format='g', digits=2)) ~ mu * 'mol mg'^{-1} ~ 'min'^{-1})), 
            size=2,
            hjust=0,
            vjust=0) +
  
    annotate("text", 
             x = min(in_data$final_substrate_conc) + max(in_data$final_substrate_conc)*0.2, 
             y = min(in_data$activity_moles_protein) + max(in_data$activity_moles_protein)*0.15, 
             label = as.expression(bquote(italic(K)[M] ~ '=' ~ .(formatC(km, format='g', digits=2)) ~ 'mM')), 
             size=2,
             hjust=0,
             vjust=0) + 
  
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme))


p1

# save the figure
fname = paste(FINAL, THIS_PROJECT_FOLDER, sprintf('%s_flex_six_plate_%s_array_%s_Km_plot.pdf', gsub(' ', '_', unique(all.df$assay_enzyme)), PLATE_NUMBER, ARRAY_NUMBER), sep='/')
ggsave(filename=fname, plot=p1, width = 65*length(unique(sample.df$sample_substrate)), height = 50*length(unique(assay.df$assay_enzyme)), units=c('mm'))


# save the km and kcat

```


