---
output:
  html_document: default
  pdf_document: default
---

---
title: "First test of D4MUV9 in lysate"
output: html_notebook
---

Copyright (C) 2018  Martin Engqvist Lab




## Experiment
This test is identical to the LOX test done on array 2, except that this time we use E. coli lysate containing a recombinant lactate oxidase. This is the form in which we would be performing the enzyme assays in an application. We test BSA and Tween/Triton as well as different substrates, just like in the other test. The enzyme in question is D4MUV9.

Experimental data collected on 301018 (DDMMYY) in Petter Wolls group at Karolinska using a BioMark instrument.


```{r setup}

#make sure the packages are installed
wants <- c('dplyr', 'reshape2', 'ggplot2')
has   <- wants %in% rownames(installed.packages())
if(any(!has)) install.packages(wants[!has], repos="http://ftp.acc.umu.se/mirror/CRAN/")

#load all the packages
lapply(wants, require, character.only = TRUE)


#find the "code" directory, this should be one folder away from the project root
while (basename(getwd()) != 'code'){
  setwd("../")
  if (getwd() == '/'){
    quit(status = 1)
  }
}

#go one folder up and set as project root directory
setwd("../")
PROJ <- getwd()
setwd(paste0(PROJ, '/data/'))

#set folder paths relative to the project root
CURRENT <- getwd()
DATA <- getwd()
RAW_EXTERNAL <- paste0(DATA, '/raw_external/')
RAW_INTERNAL <- paste0(DATA, '/raw_internal/')
INTERMEDIATE = paste0(DATA, '/intermediate/')
FINAL = paste0(DATA, '/final/')

RESULTS <- paste0(PROJ, '/results/')
FIGURES <- paste0(RESULTS, 'figures/')
PICTURES <- paste0(RESULTS, 'pictures/')

```


## Setup
Let's input the experimental paramenters for sample and assay sides

```{r}
# Test 2 block1

assay.df <- data.frame('assay_well'=c('A01','A02','A03','A04','A05','A06','A07','A08','A09','A10','A11','A12'),
                       'assay_tween'=c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1),
                        'assay_tween_unit'=c('% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)'),
                       'assay_triton_conc'=c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1),
                       'assay_triton_unit'=c('% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)'),
                       'assay_fluorescein_conc'=c(1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0, 1.0),
                       'assay_fluorescein_unit'=c('uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM'),
                       'assay_amplex_red_conc'=c(500.0, 500.0, 500.0, 500.0, 500.0, 500.0, 500.0, 500.0, 500.0, 500.0, 500.0, 500.0),
                       'assay_amplex_red_unit'=c('uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM'),                                 
                       'assay_substrate_name'=c('none', 'glycolate', 'lactate', 'none', 'glycolate', 'lactate', 'none', 'glycolate', 'lactate', 'none', 'glycolate', 'lactate'),
                       'assay_substrate_conc'=c(0.0, 10.0, 10.0, 0.0, 10.0, 10.0, 0.0, 10.0, 10.0, 0.0, 10.0, 10.0),
                       'assay_substrate_unit'=c('mM', 'mM', 'mM', 'mM', 'mM', 'mM', 'mM', 'mM', 'mM', 'mM', 'mM', 'mM'),
                        'assay_bsa_conc'=c(0.0, 0.0, 0.0, 0.0, 0.0, 0.0, 0.2, 0.2, 0.2, 0.2, 0.2, 0.2),
                       'assay_bsa_unit'=c('mg/mL', 'mg/mL', 'mg/mL', 'mg/mL', 'mg/mL', 'mg/mL', 'mg/mL', 'mg/mL', 'mg/mL', 'mg/mL', 'mg/mL', 'mg/mL'))


sample.df <- data.frame('sample_well'=c('S01','S02','S03','S04','S05','S06','S07','S08','S09','S10','S11','S12'),
                        'sample_tween_conc'=c(0.01, 0.01, 0.01, 0.01, 0.01, 0.01, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1),
                        'sample_tween_unit'=c('% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)'),
                        'sample_triton_conc'=c(0.01, 0.01, 0.01, 0.01, 0.01, 0.01, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1),
                        'sample_triton_unit'=c('% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)', '% (v/v)'),                        
                        'sample_fluorescein_conc'=c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.1),
                        'sample_fluorescein_unit'=c('uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM', 'uM'),         
                        'sample_enzyme_name'=c('D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9', 'D4MUV9'),
                        'sample_enzyme_conc'=c(0.0, 0.01, 0.02, 0.05, 0.1, 0.5, 0.0, 0.01, 0.02, 0.05, 0.1, 0.5),
                        'sample_enzyme_units'=c('relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution', 'relative_dilution'))


head(assay.df)
head(sample.df)
```


```{r}
# make all combinations of wells
mixed.df <- expand.grid(sample.df$sample_well, assay.df$assay_well)

# fix the colnames
colnames(mixed.df) <- c('sample_well', 'assay_well')

# add in the original data
mixed.df <- merge(mixed.df, sample.df, by='sample_well')
mixed.df <- merge(mixed.df, assay.df, by='assay_well')


# calculate final 
mixed.df$final_well <- paste(mixed.df$sample_well, mixed.df$assay_well, sep='-')

mixed.df$final_fluorescein_conc <- mixed.df$assay_fluorescein_conc/10 + mixed.df$sample_fluorescein_conc/1.1
mixed.df$final_fluorescein_unit <- mixed.df$assay_fluorescein_unit

mixed.df$final_amplex_red_conc <- mixed.df$assay_amplex_red_conc/10 
mixed.df$final_amplex_red_unit <- mixed.df$assay_amplex_red_unit

mixed.df$final_substrate_name <- mixed.df$assay_substrate_name
mixed.df$final_substrate_conc <- mixed.df$assay_substrate_conc/10 
mixed.df$final_substrate_unit <- mixed.df$assay_substrate_unit

mixed.df$final_bsa_conc <- mixed.df$assay_bsa_conc/10 
mixed.df$final_bsa_unit <- mixed.df$assay_bsa_unit

mixed.df$final_enzyme_name <- mixed.df$sample_enzyme_name
mixed.df$final_enzyme_conc <- mixed.df$sample_enzyme_conc/1.1 
mixed.df$final_enzyme_unit <- mixed.df$sample_enzyme_unit

head(mixed.df)
tail(mixed.df)
```




## Load data
So now we have the conditions in the final wells. Time to load up the measured data.
From the data file that goes with this script (1_1530021086_array_4_raw.csv) I manually exported the ROX and FAM signal separately to tsv files. Load the data and then combine with my condition file.

```{r}
# read the experimental data obtained from the BioMark
rox.df <- read.table(paste(INTERMEDIATE, 'flex_six_first_test', 'Array_4', '3_1530021086_array_4_rox.tsv', sep='/'), sep='\t', header=TRUE)
colnames(rox.df) <- c(c('final_well'), paste0('rox_cycle_', seq(length(colnames(rox.df))-1)))

fam.df <- read.table(paste(INTERMEDIATE, 'flex_six_first_test', 'Array_4', '4_1530021086_array_4_fam.tsv', sep='/'), sep='\t', header=TRUE)
colnames(fam.df) <- c(c('final_well'), paste0('fam_cycle_', seq(length(colnames(fam.df))-1)))


# Add the data into my frame with concentration info
all.df <- merge(mixed.df, rox.df, by='final_well')
all.df <- merge(all.df, fam.df, by='final_well')

head(all.df)
```


Looks good! Let's export this to a flatfile for future use.


```{r}
write.table(all.df, file=paste(FINAL, 'flex_six_first_test', 'array_4_all_data_combined.tsv', sep='/'), sep='\t', row.names=FALSE)
```



## Plotting results
Time to plot!


A look at the raw data as a heatmap.

```{r}


# ----- Define a function for plotting a matrix ----- #
# from http://www.phaget4.org/R/image_matrix.html
myImagePlot <- function(x, ...){
     min <- 0
     max <- 65000
     yLabels <- rownames(x)
     xLabels <- colnames(x)
     title <-c()
  # check for additional function arguments
  if( length(list(...)) ){
    Lst <- list(...)
    if( !is.null(Lst$zlim) ){
       min <- Lst$zlim[1]
       max <- Lst$zlim[2]
    }
    if( !is.null(Lst$yLabels) ){
       yLabels <- c(Lst$yLabels)
    }
    if( !is.null(Lst$xLabels) ){
       xLabels <- c(Lst$xLabels)
    }
    if( !is.null(Lst$title) ){
       title <- Lst$title
    }
  }
# check for null values
if( is.null(xLabels) ){
   xLabels <- c(1:ncol(x))
}
if( is.null(yLabels) ){
   yLabels <- c(1:nrow(x))
}

layout(matrix(data=c(1,2), nrow=1, ncol=2), widths=c(4,1), heights=c(1,1))

 # Red and green range from 0 to 1 while Blue ranges from 1 to 0
 ColorRamp <- colorRampPalette(c('#222222', '#EEEEEE'))(n = 256)
 ColorLevels <- seq(min, max, length=length(ColorRamp))

 # Reverse Y axis
 reverse <- nrow(x) : 1
 yLabels <- yLabels[reverse]
 x <- x[reverse,]

 # Data Map
 par(mar = c(3,5,2.5,2))
 image(1:length(xLabels), 1:length(yLabels), t(x), col=ColorRamp, xlab="",
 ylab="", axes=FALSE, zlim=c(min,max))
 if( !is.null(title) ){
    title(main=title)
 }
axis(BELOW<-1, at=1:length(xLabels), labels=xLabels, cex.axis=0.7)
 axis(LEFT <-2, at=1:length(yLabels), labels=yLabels, las= HORIZONTAL<-1,
 cex.axis=0.7)

 # Color Scale
 par(mar = c(3,2.5,2.5,2))
 image(1, ColorLevels,
      matrix(data=ColorLevels, ncol=length(ColorLevels),nrow=1),
      col=ColorRamp,
      xlab="",ylab="",
      xaxt="n")

 layout(1)
}
# ----- END plot function ----- #



make_heatmap <- function(mat_data){

  color_scale <- colorRampPalette(c('#1A1A1A', '#EEEEEE'))(n = 256)
  
  # adjust the color scale. Max value sems to be around 65,000
  max_val <- max(mat_data)
  if (max_val > 65000){
    color_end <- 256
  }else{
    color_end <- round((max_val/65000) * 256)
  }
  color_scale <- color_scale[1:color_end]
  
  image(mat_data, 
          axes=FALSE, 
          col = color_scale, 
          xlab='Assay', 
          ylab='Sample')
}


# make a subset of the data frames for plotting
rox_data.df <- all.df[, which(names(all.df) %in% c('assay_well', 'sample_well', 'rox_cycle_15'))]
fam_data.df <- all.df[, which(names(all.df) %in% c('assay_well', 'sample_well', 'fam_cycle_15'))]

# make wide format
rox_data_wide.df <- dcast(rox_data.df, assay_well ~ sample_well, value.var='rox_cycle_15')
fam_data_wide.df <- dcast(fam_data.df, assay_well ~ sample_well, value.var='fam_cycle_15')

# use first column as row names
rox.mat_data <- data.matrix(rox_data_wide.df[, 2:length(names(rox_data_wide.df))]) 
rnames <- rox_data_wide.df[,'assay_well']   # assign labels in the "File" column to "rnames"
rownames(rox.mat_data) <- rnames 

fam.mat_data <- data.matrix(fam_data_wide.df[, 2:length(names(fam_data_wide.df))]) 
rnames <- fam_data_wide.df[,'assay_well']   # assign labels in the "File" column to "rnames"
rownames(fam.mat_data) <- rnames 


# Now make the heatmap
myImagePlot(rox.mat_data, title=('ROX channel'))
myImagePlot(fam.mat_data, title=('FAM channel'))
```





Let's first just have a look at all the ROX and FAM channel data.

```{r}
plot(all.df$rox_cycle_15,
     ylab='Signal in the ROX channel (fluorescence)')

plot(all.df$fam_cycle_15,
     ylab='Signal in the FAM channel (fluorescence)')
```



First, let's see what the fluorescein loading control looks like.

```{r}
hist(all.df[all.df$sample_tween_conc==0.1, ]$fam_cycle_15, 
     main='Loading control visualzation for the Tween/Triton 0.1 condition',
     xlab='Signal in the FAM channel (fluorescence)')

hist(all.df[all.df$sample_tween_conc==0.01, ]$fam_cycle_15, 
     main='Loading control visualzation for the Tween/Triton 0.01 condition',
     xlab='Signal in the FAM channel (fluorescence)')

```

This looks good. Any bad loading should lead to half the signal. I don't seem to see that. So both tween/triton concentrations seem to work well to load E. coli lysate.


Next let's compare whether the sample tween/triton concentration had an effect on the enzyme activity.

```{r}
ggplot(all.df, aes(x=final_enzyme_conc, y=rox_cycle_15, shape = factor(sample_tween_conc), color = factor(sample_tween_conc)))+
    geom_point() + 
    scale_color_manual(values = c('0.01' = 'darkorange2', '0.1' = 'darkgreen'), name='Tween and Triton conc.') + 
    scale_shape_manual(values = c('0.01' = 21, '0.1' = 25), name='Tween and Triton conc.') +
    ggtitle('D4MUV9 dilution series in different detergents, activity with three substrates') +
    xlab('Relative enzyme dilution') + 
    ylab('Signal in ROX channel (fluorescence)') +
    facet_grid(~final_substrate_name)

```

The detergent seems to have very little effect on the activity. If anything, the 0.1 concentration appears to give more variable data for the glycolate measurements. Oddly though there seems to be a falloff in activity at the lowest dilution 2x dilution (the 0.5 value). This is particularly visible for the lactate signal. So somwhat higher dilution rates appear to be optimal to detect activity. About 10x dilution or so.


```{r}
ggplot(all.df, aes(x=final_enzyme_conc, y=rox_cycle_15, shape = factor(final_bsa_conc), color = factor(final_bsa_conc)))+
    geom_point() + 
    scale_color_manual(values = c('0' = 'darkorange2', '0.02' = 'darkgreen'), name='BSA conc.') + 
    scale_shape_manual(values = c('0' = 21, '0.02' = 25), name='BSA conc.') +
    ggtitle('D4MUV9 dilution series in different detergents, activity with three substrates') +
    xlab('Final enzyme concentration (mU/mL)') + 
    ylab('Signal in ROX channel (fluorescence)') +
    facet_grid(~final_substrate_name)
```



As previously seen with LOX, the BSA has little to no effect on the assays (compare orange and green). So peroxidase adsorption is not a problem it seems.





We did do "kinetic" measurements of the enzymes. Meeaning that we cycled 15 times at 20C for 1 minute. That gives us 15 measurements of fluorescence. I'm interested in seeing whether we can actually see signal increase over time.



```{r}

# subset the data
rox_data.df <- all.df[, which(names(all.df) %in% c('assay_substrate_name', 'final_bsa_conc', 'sample_tween_conc', 'sample_enzyme_conc', 'rox_cycle_1', 'rox_cycle_2', 'rox_cycle_3', 'rox_cycle_4', 'rox_cycle_5', 'rox_cycle_6', 'rox_cycle_7', 'rox_cycle_8', 'rox_cycle_9', 'rox_cycle_10', 'rox_cycle_11', 'rox_cycle_12', 'rox_cycle_13', 'rox_cycle_14', 'rox_cycle_15'))]
fam_data.df <- all.df[, which(names(all.df) %in% c('assay_substrate_name', 'final_bsa_conc', 'sample_tween_conc', 'sample_enzyme_conc', 'fam_cycle_1', 'fam_cycle_2', 'fam_cycle_3', 'fam_cycle_4', 'fam_cycle_5', 'fam_cycle_6', 'fam_cycle_7', 'fam_cycle_8', 'fam_cycle_9', 'fam_cycle_10', 'fam_cycle_11', 'fam_cycle_12', 'fam_cycle_13', 'fam_cycle_14', 'fam_cycle_15'))]


# I need to convert the data from wide format to long format
rox_data_long.df <- melt(rox_data.df, id=c('assay_substrate_name', 'final_bsa_conc', 'sample_tween_conc', 'sample_enzyme_conc')) 
fam_data_long.df <- melt(fam_data.df, id=c('assay_substrate_name', 'final_bsa_conc', 'sample_tween_conc', 'sample_enzyme_conc')) 

# Convert the cycle number from text to ints
rox_data_long_cycles.df <- rox_data_long.df
rox_data_long_cycles.df$variable <- as.numeric(gsub("rox_cycle_", '', rox_data_long.df$variable) )

fam_data_long_cycles.df <- fam_data_long.df
fam_data_long_cycles.df$variable <- as.numeric(gsub("fam_cycle_", '', fam_data_long.df$variable) )
 

# Done. Ready to plot now
ggplot(rox_data_long_cycles.df[rox_data_long_cycles.df$sample_tween_conc == 0.01, ], aes(x=variable, y=value, shape = factor(sample_enzyme_conc), color = factor(sample_enzyme_conc)))+
    geom_point() + 
    guides(color=guide_legend(title="Enzyme rel. conc.")) +
    guides(shape=guide_legend(title="Enzyme rel. conc.")) +
    ggtitle('A9QH69 dilution series in Tween/Triton 0.01, activity with three substrates') +
    xlab('Cycle (minutes)') + 
    ylab('Signal in ROX channel (fluorescence)') +
    facet_grid(~assay_substrate_name)

ggplot(rox_data_long_cycles.df[rox_data_long_cycles.df$sample_tween_conc == 0.1, ], aes(x=variable, y=value, shape = factor(sample_enzyme_conc), color = factor(sample_enzyme_conc)))+
    geom_point() + 
    guides(color=guide_legend(title="Enzyme rel. conc.")) +
    guides(shape=guide_legend(title="Enzyme rel. conc.")) +
    ggtitle('A9QH69 dilution series in Tween/Triton 0.1, activity with three substrates') +
    xlab('Cycle (minutes)') + 
    ylab('Signal in ROX channel (fluorescence)') +
    facet_grid(~assay_substrate_name)

```


The reactions are already finished for the less dilute samples when we put the chip in the BioMark. The 100x dilutet sample (relative concentration 0.01) does look good, however. The least dilute sample (2x dilution, 0.5 relative concentration) performs more poorly than less dilute samples.

