---
output:
  html_document: default
  pdf_document: default
---

---
title: "Resorufin standard curve, measured in Biomark"
output: html_notebook
---

Copyright (C) 2020  Martin Engqvist Lab




## Experiment
.....

Experimental data collected on 160719 (DDMMYY) in Petter Wolls group at Karolinska using a BioMark instrument.


```{r setup}

#make sure the packages are installed
wants <- c('dplyr', 'reshape2', 'ggplot2', 'tidyverse', 'assertthat')
has   <- wants %in% rownames(installed.packages())
if(any(!has)) install.packages(wants[!has], repos="http://ftp.acc.umu.se/mirror/CRAN/")

#load all the packages
lapply(wants, require, character.only = TRUE)


#find the "code" directory, this should be one folder away from the project root
while (basename(getwd()) != 'code'){
  setwd("../")
  if (getwd() == '/'){
    quit(status = 1)
  }
}

#go one folder up and set as project root directory
setwd("../")
PROJ <- getwd()
setwd(paste0(PROJ, '/data/'))

#set folder paths relative to the project root
CURRENT <- getwd()
DATA <- getwd()
RAW_EXTERNAL <- paste0(DATA, '/raw_external/')
RAW_INTERNAL <- paste0(DATA, '/raw_internal/')
INTERMEDIATE = paste0(DATA, '/intermediate/')
FINAL = paste0(DATA, '/final/')

RESULTS <- paste0(PROJ, '/results/')
FIGURES <- paste0(RESULTS, 'figures/')
PICTURES <- paste0(RESULTS, 'pictures/')

THIS_PROJECT_FOLDER <- 'resorufin_std_curves'

# load helper functions
source(paste0(PROJ, '/code/', 'flex_six_helper_functions.R'))
print(paste0('helper script version: ', VERSION))

```



Setup plotting preferences

```{r}
font_size = 7
plt_height = 65
plt_width = 89
line_size = 0.5
line_width <- 0.5
error_size = 0.25
pd = position_dodge(0.9)

gen_col = '#141A46'
nat_col = '#EC8B5E'
neg_col = '#868686'



ggplot_theme <- theme(text = element_text( size=font_size), 
                      axis.text.x = element_text(angle = 0, hjust = 1, size=font_size),
                      axis.text.y = element_text(size=font_size),
                      axis.line = element_line(color = "black", size=line_size),
                      axis.ticks.x = element_line(color = 'black', size=line_size),
                      axis.ticks.y = element_line(color = 'black', size=line_size),
                      #legend.text = element_text( size=font_size),
                      legend.position = c(0, 1),
                      legend.justification = c(0, 1),
                      legend.background = element_blank(),
                      legend.direction = 'horizontal',
                      legend.text=element_text(size=font_size),
                      plot.title = element_text( size=font_size),
                      panel.background = element_blank(),
                      panel.border = element_rect(color = 'black', size=line_size),
                      #panel.margin = 
                      panel.grid.major = element_blank(), 
                      panel.grid.minor = element_blank(),
                      panel.grid.major.x = element_blank(), 
                      panel.grid.major.y = element_blank(), 
                      panel.grid.minor.x = element_blank(), 
                      panel.grid.minor.y = element_blank() 
)
```



```{r}
load_data <- function(chip_number, array_number, rox_data, fam_data, repeat_num, experiment){
  
  # load sample layout
  assay.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, sprintf('chip%s_array%s_assay.tsv', chip_number, array_number), sep='/'), sep='\t', header=TRUE)
  
  # load array layout
  sample.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, sprintf('chip%s_array%s_sample.tsv', chip_number, array_number), sep='/'), sep='\t', header=TRUE)
  
  # mix the two
  mixed.df <- mix_wells(assay.df, sample.df)
  
  # read the experimental data obtained from the BioMark
  rox.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, rox_data, sep='/'), sep='\t', header=TRUE)
  fam.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, fam_data, sep='/'), sep='\t', header=TRUE)
  
  # convert to long format
  all.df <- combine_data(mixed.df, rox.df, fam.df)
  
  # specify experiment-specific parameters
  all.df$repeat_num <- repeat_num
  all.df$experiment <- experiment
  
  return(all.df)
}
```





## Load all the data
```{r}

data1.df <- load_data(chip_number=3, array_number=4, 
                     rox_data='1530024086_chip_3_array_4_rox.tsv', 
                     fam_data='1530024086_chip_3_array_4_fam.tsv', 
                     repeat_num=1, 
                     experiment='low_range')

data2.df <- load_data(chip_number=3, array_number=5, 
                     rox_data='1530024086_chip_3_array_5_rox.tsv', 
                     fam_data='1530024086_chip_3_array_5_fam.tsv', 
                     repeat_num=1, 
                     experiment='high_range')

data3.df <- load_data(chip_number=3, array_number=6, 
                     rox_data='1530024086_chip_3_array_6_rox.tsv', 
                     fam_data='1530024086_chip_3_array_6_fam.tsv', 
                     repeat_num=2, 
                     experiment='low_range')

data4.df <- load_data(chip_number=4, array_number=4, 
                     rox_data='1530024197_chip_4_array_4_rox.tsv', 
                     fam_data='1530024197_chip_4_array_4_fam.tsv', 
                     repeat_num=2, 
                     experiment='high_range')

all.df <- rbind(data1.df, data2.df, data3.df, data4.df)

head(all.df)
tail(all.df)

# Let's export this to a flatfile for future use.
write.table(all.df, file=paste(FINAL, THIS_PROJECT_FOLDER, 'flex_six_std_curves_all_data_combined.tsv', sep='/'), sep='\t', row.names=FALSE)

```



## Plotting results
Time to plot!


Plot rox and fam values of first and last cycle

```{r}

# low range, first experiment
plot_raw_data_heatmaps(all.df[all.df$experiment=='low_range' & all.df$repeat_num==1, ], min=0, max=65000)

# low range, second experiment
plot_raw_data_heatmaps(all.df[all.df$experiment=='low_range' & all.df$repeat_num==2, ], min=0, max=65000)

# high range, first experiment
plot_raw_data_heatmaps(all.df[all.df$experiment=='high_range' & all.df$repeat_num==1, ], min=0, max=65000)

# high range, second experiment
plot_raw_data_heatmaps(all.df[all.df$experiment=='high_range' & all.df$repeat_num==2, ], min=0, max=65000)
```




First we plot the raw values observed in all the wells

```{r}
ggplot(all.df[all.df$cycle==1, ], aes(x=final_substrate_conc, y=rox_value, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Raw data at cycle 1') +
    xlab('Concentration') + 
    ylab('Signal') +
    facet_grid(cols=vars(experiment), rows=vars(repeat_num))

```
Concentrations at 10 uM and above are very clearly too high. The curve is not linear any more. Let's remove value above and re-plot.



Remove the high range data since this reaches a chipau

```{r}
all.df.subset <- all.df[all.df$final_substrate_conc <= 5, ]

ggplot(all.df.subset[all.df.subset$cycle==1, ], aes(x=final_substrate_conc, y=rox_value, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Raw data at cycle 1') +
    xlab('Concentration') + 
    ylab('Signal') +
    facet_grid(cols=vars(experiment), rows=vars(repeat_num))

```
This looks much better! Strangly the high_range samples have a large spread, but the low_range samples look good.




I need to subtract the base fluorescence (the 0 sample) from these.
```{r}

# first get the values from control
control.df <- all.df.subset[all.df.subset$final_substrate_conc==0, c('cycle', 'repeat_num', 'rox_value', 'experiment')]

# rename the rox value
colnames(control.df)[colnames(control.df) == 'rox_value'] <- 'rox_value_control'

# do a left join to add on the controls to the samples
df.combined <- merge(all.df.subset, control.df, by=(c('cycle', 'experiment', 'repeat_num')), all.x=TRUE)

# now calculate the normalized values
df.combined$rox_value_norm <- df.combined$rox_value - df.combined$rox_value_control

# remove the control samples, they are not needed anymore as the data is now normalized
df.combined <- df.combined[!df.combined$final_substrate_conc==0, ]
df.combined$final_substrate_norm_unit <- unique(df.combined$final_substrate_unit)

# drop unneeded columns
all.df.subset.norm <- df.combined[, -which(names(df.combined) %in% c('rox_value_control'))]

# re-order the data
all.df.subset.norm <- all.df.subset.norm %>% 
  select(cycle, assay_well, sample_well, assay_enzyme, final_enzyme_conc, final_enzyme_unit, sample_substrate, final_substrate_conc, final_substrate_unit, repeat_num, rox_value_norm, final_substrate_norm_unit, experiment)



ggplot(all.df.subset.norm[all.df.subset.norm$cycle==1, ], aes(x=final_substrate_conc, y=rox_value_norm, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Background-subtracted data at cycle 1') +
    xlab('Concentration') + 
    ylab('Signal') +
    facet_grid(cols=vars(experiment), rows=vars(repeat_num))
```



Fit a linear regression for all the data together

```{r}
in_data <- all.df.subset.norm[all.df.subset.norm$cycle==1 & all.df.subset.norm$experiment == 'low_range' & all.df.subset.norm$repeat_num == 1, ]

# Fit a linear regression to get the slope.
# make a model for only repeat 1 of the low range
model <- lm(in_data$rox_value_norm ~ in_data$final_substrate_conc)

rep1_intercept <- coef(model)[[1]]
rep1_slope <- coef(model)[[2]]
rep1_fit <- summary(model)$r.squared

#model
print('Repeat 1, low range')
sprintf('Intercept: %s', rep1_intercept)
sprintf('Slope: %s', rep1_slope)
sprintf('R2: %s', rep1_fit)
print('')

# plot the graph
p1 <- ggplot(in_data, aes(x=final_substrate_conc, y=rox_value_norm)) +
    theme_bw() +
    ggplot_theme +
    geom_smooth(method = 'lm', se = FALSE) +
    geom_point(cex=1.0, alpha=0.5, shape=16) + 
    ggtitle('Fluidigm') +
    xlab(expression(paste('Resorufin (', mu, 'M', ')') ) ) + 
    ylab('Fluorescence') + 
    annotate("text", 
            x = 0, 
            y = max(in_data$rox_value_norm)*0.95, 
            label = as.expression(bquote(italic(R)^{2} ~ '=' ~ .(formatC(rep1_fit, format='g', digits=3)))), 
            size=2,
            hjust=0,
            vjust=0) + 
  
    annotate("text", 
            x = 0, 
            y = max(in_data$rox_value_norm)*0.85, 
            label = as.expression(bquote('y' ~ '=' ~ .(formatC(rep1_slope, format='g', digits=3)) ~ 'x + ' ~ .(formatC(rep1_intercept, format='g', digits=3)))), 
            size=2,
            hjust=0,
            vjust=0)

p1

# save the figure
fname = paste(FIGURES, sprintf('resorufin_standard_curve_biomark.pdf'), sep='/')
ggsave(filename=fname, plot=p1, width = 65, height = 50, units=c('mm'))
```




Print out the final data

```{r}
print('Standard curve according to: y = kx + m')
sprintf('y = %sx + %s    with an R2 of %s', round(rep1_slope), round(rep1_intercept), round(rep1_fit, 3))
```


