---
output:
  html_document: default
  pdf_document: default
---

---
title: "Resorufin standard curve, measured in microtiter plate"
output: html_notebook
---

Copyright (C) 2020  Martin Engqvist Lab




## Experiment
.....



```{r setup}

#make sure the packages are installed
wants <- c('dplyr', 'reshape2', 'ggplot2', 'tidyverse', 'assertthat')
has   <- wants %in% rownames(installed.packages())
if(any(!has)) install.packages(wants[!has], repos="http://ftp.acc.umu.se/mirror/CRAN/")

#load all the packages
lapply(wants, require, character.only = TRUE)


#find the "code" directory, this should be one folder away from the project root
while (basename(getwd()) != 'code'){
  setwd("../")
  if (getwd() == '/'){
    quit(status = 1)
  }
}

#go one folder up and set as project root directory
setwd("../")
PROJ <- getwd()
setwd(paste0(PROJ, '/data/'))

#set folder paths relative to the project root
CURRENT <- getwd()
DATA <- getwd()
RAW_EXTERNAL <- paste0(DATA, '/raw_external/')
RAW_INTERNAL <- paste0(DATA, '/raw_internal/')
INTERMEDIATE = paste0(DATA, '/intermediate/')
FINAL = paste0(DATA, '/final/')

RESULTS <- paste0(PROJ, '/results/')
FIGURES <- paste0(RESULTS, 'figures/')
PICTURES <- paste0(RESULTS, 'pictures/')

THIS_PROJECT_FOLDER <- 'resorufin_std_curves'

```



Setup plotting preferences

```{r}
font_size = 7
plt_height = 65
plt_width = 89
line_size = 0.5
line_width <- 0.5
error_size = 0.25
pd = position_dodge(0.9)

gen_col = '#141A46'
nat_col = '#EC8B5E'
neg_col = '#868686'



ggplot_theme <- theme(text = element_text( size=font_size), 
                      axis.text.x = element_text(angle = 0, hjust = 1, size=font_size),
                      axis.text.y = element_text(size=font_size),
                      axis.line = element_line(color = "black", size=line_size),
                      axis.ticks.x = element_line(color = 'black', size=line_size),
                      axis.ticks.y = element_line(color = 'black', size=line_size),
                      #legend.text = element_text( size=font_size),
                      legend.position = c(0, 1),
                      legend.justification = c(0, 1),
                      legend.background = element_blank(),
                      legend.direction = 'horizontal',
                      legend.text=element_text(size=font_size),
                      plot.title = element_text( size=font_size),
                      panel.background = element_blank(),
                      panel.border = element_rect(color = 'black', size=line_size),
                      #panel.margin = 
                      panel.grid.major = element_blank(), 
                      panel.grid.minor = element_blank(),
                      panel.grid.major.x = element_blank(), 
                      panel.grid.major.y = element_blank(), 
                      panel.grid.minor.x = element_blank(), 
                      panel.grid.minor.y = element_blank() 
)
```


## Load the data
```{r}

data.df <- read.table(paste(RAW_INTERNAL, THIS_PROJECT_FOLDER, 'normal_plate_std_curve.tsv', sep='/'), sep='\t', header=TRUE)

head(data.df)
tail(data.df)


```



Convert data to long format

```{r}
data.df.long <- melt(data.df, id=c('resorufin_conc', 'units')) 
colnames(data.df.long) <- c('resorufin_conc', 'resorufin_conc_unit', 'repeat_num', 'fluorescence')

# convert repeat to numeric
data.df.long$repeat_num <- as.integer(gsub('rep', '', data.df.long$repeat_num))

# and fluorescence to double
data.df.long$fluorescence <- as.numeric(data.df.long$fluorescence)

head(data.df.long)
tail(data.df.long)
```



## Plotting results
Time to plot!


First we plot the raw values observed in all the wells

```{r}
in_data <- data.df.long
ggplot(in_data, aes(x=resorufin_conc, y=fluorescence, color=factor(resorufin_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Raw data') +
    xlab('Concentration (uM)') + 
    ylab('Fluorescence') +
    facet_grid(repeat_num ~ .)

```
Concentrations above 5 uM are very clearly too high. The curve is not linear any more. Let's remove value above and re-plot.



Remove the high range data since this reaches a plateau

```{r}
all.df.subset <- data.df.long[data.df.long$resorufin_conc <= 5, ]

in_data <- all.df.subset
ggplot(in_data, aes(x=resorufin_conc, y=fluorescence, color=factor(resorufin_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Raw data') +
    xlab('Concentration (uM)') + 
    ylab('Fluorescence') +
    facet_grid(repeat_num ~ .)

```
This looks good!



Let's look at the low range values, just to see what's going on.

```{r}

in_data <- all.df.subset[all.df.subset$fluorescence < 2500, ]
ggplot(in_data, aes(x=resorufin_conc, y=fluorescence, color=factor(resorufin_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Raw data') +
    xlab('Concentration (uM)') + 
    ylab('Fluorescence') +
    facet_grid(repeat_num ~ .)
```
Looks nice and tidy!





Fit a linear regression for all the data together

```{r}

# plot with a linear regression

in_data <- all.df.subset

# Fit a linear regression to get the slope.
# make a model for only repeat 1 of the low range
model <- lm(in_data$fluorescence ~ in_data$resorufin_conc)

all_intercept <- coef(model)[[1]]
all_slope <- coef(model)[[2]]
all_fit <- summary(model)$r.squared

#model
print('Repeat 1, low range')
sprintf('Intercept: %s', all_intercept)
sprintf('Slope: %s', all_slope)
sprintf('R2: %s', all_fit)
print('')

# plot the graph
p1 <- ggplot(in_data, aes(x=resorufin_conc, y=fluorescence)) +
    theme_bw() +
    ggplot_theme +
    geom_smooth(method = 'lm', se = FALSE) +
    geom_point(cex=1.0, alpha=0.5, shape=16) + 
    ggtitle('Plate reader') +
    xlab(expression(paste('Resorufin (', mu, 'M', ')') ) ) + 
    ylab('Fluorescence') + 
    annotate("text", 
            x = 0, 
            y = max(in_data$fluorescence)*0.95, 
            label = as.expression(bquote(italic(R)^{2} ~ '=' ~ .(formatC(all_fit, format='g', digits=3)))), 
            size=2,
            hjust=0,
            vjust=0) + 
  
    annotate("text", 
            x = 0, 
            y = max(in_data$fluorescence)*0.85, 
            label = as.expression(bquote('y' ~ '=' ~ .(formatC(all_slope, format='g', digits=3)) ~ 'x + ' ~ .(formatC(all_intercept, format='g', digits=3)))), 
            size=2,
            hjust=0,
            vjust=0)

p1

# save the figure
fname = paste(FIGURES, sprintf('resorufin_standard_curve_plate_reader.pdf'), sep='/')
ggsave(filename=fname, plot=p1, width = 65, height = 50, units=c('mm'))
```





Print out the parameters

```{r}
print('Standard curve according to: y = kx + m')
sprintf('y = %sx + %s    with an R2 of %s', round(all_slope), round(all_intercept), round(all_fit, 3))
```


