---
output:
  html_document: default
  pdf_document: default
---

---
title: "Test of whether kinetic constants can be obtained with BioMark using the 96x96 chip"
output: html_notebook
---

Copyright (C) 2019  Martin Engqvist Lab




## Experiment
.....

Experimental data collected on xxxxxx (DDMMYY) in Petter Wolls group at Karolinska using a BioMark instrument.

Assay conditions, oxidases:


20mM HEPES, pH 7.4

50uM Amplex Red

0,1U/ml HRP

0,1uM fluorescein

0,01mg/ml BSA

0,01% Triton-X-100 and Tween20

1x enzyme concentration: 1,5mg/ml

```{r setup}

#make sure the packages are installed
wants <- c('dplyr', 'reshape2', 'ggplot2', 'tidyverse')
has   <- wants %in% rownames(installed.packages())
if(any(!has)) install.packages(wants[!has], repos="http://ftp.acc.umu.se/mirror/CRAN/")

#load all the packages
lapply(wants, require, character.only = TRUE)


#find the "code" directory, this should be one folder away from the project root
while (basename(getwd()) != 'code'){
  setwd("../")
  if (getwd() == '/'){
    quit(status = 1)
  }
}

#go one folder up and set as project root directory
setwd("../")
PROJ <- getwd()
setwd(paste0(PROJ, '/data/'))

#set folder paths relative to the project root
CURRENT <- getwd()
DATA <- getwd()
RAW_EXTERNAL <- paste0(DATA, '/raw_external/')
RAW_INTERNAL <- paste0(DATA, '/raw_internal/')
INTERMEDIATE = paste0(DATA, '/intermediate/')
FINAL = paste0(DATA, '/final/')

RESULTS <- paste0(PROJ, '/results/')
FIGURES <- paste0(RESULTS, 'figures/')
PICTURES <- paste0(RESULTS, 'pictures/')

```


## Define variables specific for this plate and array
```{r}
THIS_PROJECT_FOLDER <- 'Biomark_16072019_data'
DATA_FOLDER <- '1.1.3.15_96.96'
FAM_DATA <- '96x96_assay_fam.tsv'
ROX_DATA <- '96x96_assay_rox.tsv'

R2_CUTOFF <- 0.8
TOLERANCE_AT_ZERO <- 0.05
TOLERANCE_AT_VMAX <- 0.05

```


## Setup

Let's input the experimental paramenters for sample and assay sides

```{r}
assay.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, DATA_FOLDER, '1362478106_96x96_assay_side.tsv', sep='/'), sep='\t', header=TRUE)

sample.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, DATA_FOLDER, '1362478106_96x96_sample_side.tsv', sep='/'), sep='\t', header=TRUE)


print(assay.df)
print(sample.df)
```





```{r}
# make all combinations of wells
mixed.df <- expand.grid(sample.df$sample_well, assay.df$assay_well)

# fix the colnames
colnames(mixed.df) <- c('sample_well', 'assay_well')

# add in the original data
mixed.df <- merge(mixed.df, sample.df, by='sample_well')
mixed.df <- merge(mixed.df, assay.df, by='assay_well')


# calculate final concentrations of things
mixed.df$final_well <- paste(mixed.df$sample_well, mixed.df$assay_well, sep='-')

mixed.df$final_substrate_name <- mixed.df$sample_substrate_name
mixed.df$final_substrate_conc <- mixed.df$sample_substrate_conc/1.1
mixed.df$final_substrate_unit <- mixed.df$sample_substrate_unit

mixed.df$final_enzyme_name <- mixed.df$assay_enzyme_name
mixed.df$final_enzyme_conc <- mixed.df$assay_enzyme_conc/10
mixed.df$final_enzyme_unit <- mixed.df$assay_enzyme_unit

head(mixed.df)
tail(mixed.df)
```




## Load data


```{r}
# read the experimental data obtained from the BioMark
rox.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, DATA_FOLDER, ROX_DATA, sep='/'), sep='\t', header=TRUE)
fam.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, DATA_FOLDER, FAM_DATA, sep='/'), sep='\t', header=TRUE)

# convert to long format
rox.df.long <- melt(rox.df, id=c('Chamber.ID')) 
colnames(rox.df.long) <- c('final_well', 'cycle', 'rox_value')
rox.df.long$cycle <- as.integer(gsub('X', '', rox.df.long$cycle))

fam.df.long <- melt(fam.df, id=c('Chamber.ID')) 
colnames(fam.df.long) <- c('final_well', 'cycle', 'fam_value')
fam.df.long$cycle <- as.integer(gsub('X', '', fam.df.long$cycle))


# Add the data into my frame with concentration info
all.df <- merge(fam.df.long, rox.df.long, by=c('final_well', 'cycle'))
all.df <- merge(all.df, mixed.df, by=c('final_well'))

head(all.df)
```


Let's export this to a flatfile for future use.


```{r}
write.table(all.df, file=paste(FINAL, THIS_PROJECT_FOLDER, DATA_FOLDER, '96x96_all_data_combined.tsv', sep='/'), sep='\t', row.names=FALSE)
```



Setup function for plotting heatmaps

```{r}

# ----- Define a function for plotting a matrix ----- #
# from http://www.phaget4.org/R/image_matrix.html
myImagePlot <- function(x, max, min, ...){
     yLabels <- rownames(x)
     xLabels <- colnames(x)
     title <-c()
  # check for additional function arguments
  if( length(list(...)) ){
    Lst <- list(...)
    if( !is.null(Lst$zlim) ){
       min <- Lst$zlim[1]
       max <- Lst$zlim[2]
    }
    if( !is.null(Lst$yLabels) ){
       yLabels <- c(Lst$yLabels)
    }
    if( !is.null(Lst$xLabels) ){
       xLabels <- c(Lst$xLabels)
    }
    if( !is.null(Lst$title) ){
       title <- Lst$title
    }
  }
# check for null values
if( is.null(xLabels) ){
   xLabels <- c(1:ncol(x))
}
if( is.null(yLabels) ){
   yLabels <- c(1:nrow(x))
}

layout(matrix(data=c(1,2), nrow=1, ncol=2), widths=c(4,1), heights=c(1,1))

 # Red and green range from 0 to 1 while Blue ranges from 1 to 0
 ColorRamp <- colorRampPalette(c('#222222', '#EEEEEE'))(n = 256)
 ColorLevels <- seq(min, max, length=length(ColorRamp))

 # Reverse Y axis
 reverse <- nrow(x) : 1
 yLabels <- yLabels[reverse]
 x <- x[reverse,]

 # Data Map
 par(mar = c(3,5,2.5,2))
 image(1:length(xLabels), 1:length(yLabels), t(x), col=ColorRamp, xlab="",
 ylab="", axes=FALSE, zlim=c(min,max))
 if( !is.null(title) ){
    title(main=title)
 }
axis(BELOW<-1, at=1:length(xLabels), labels=xLabels, cex.axis=0.7)
 axis(LEFT <-2, at=1:length(yLabels), labels=yLabels, las= HORIZONTAL<-1,
 cex.axis=0.7)

 # Color Scale
 par(mar = c(3,2.5,2.5,2))
 image(1, ColorLevels,
      matrix(data=ColorLevels, ncol=length(ColorLevels),nrow=1),
      col=ColorRamp,
      xlab="",ylab="",
      xaxt="n")

 layout(1)
}
# ----- END plot function ----- #
```



## Plotting results
Time to plot!


Check whether everything was loaded ok, plot values of first cycle

```{r}


# make a subset of the data frames for plotting
rox_data_first_cycle.df <- all.df[all.df$cycle == 1, which(names(all.df) %in% c('assay_well', 'sample_well', 'rox_value'))]
fam_data_first_cycle.df <- all.df[all.df$cycle == 1, which(names(all.df) %in% c('assay_well', 'sample_well', 'fam_value'))]

# make wide format
rox_data_wide.df <- dcast(rox_data_first_cycle.df, assay_well ~ sample_well, value.var='rox_value')
fam_data_wide.df <- dcast(fam_data_first_cycle.df, assay_well ~ sample_well, value.var='fam_value')

# use first column as row names
rox.mat_data <- data.matrix(rox_data_wide.df[, 2:length(names(rox_data_wide.df))]) 
rnames <- rox_data_wide.df[,'assay_well']   # assign labels in the "File" column to "rnames"
rownames(rox.mat_data) <- rnames 

fam.mat_data <- data.matrix(fam_data_wide.df[, 2:length(names(fam_data_wide.df))]) 
rnames <- fam_data_wide.df[,'assay_well']   # assign labels in the "File" column to "rnames"
rownames(fam.mat_data) <- rnames 


# Now make the heatmap
myImagePlot(rox.mat_data, max=65000, min=0, title=('ROX channel'))
myImagePlot(fam.mat_data, max=65000, min=0, title=('FAM channel'))
```



** Here I should add a step where all samples with "bad data", as indicated by no FAM fluorescence, are removed from the dataset.
```{r}

```



First we plot the raw kinetics observed in all the wells

```{r fig.height = 20, fig.width = 10}

ggplot(all.df, aes(x=cycle, y=rox_value, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Raw data measured over time') +
    xlab('Cycle') + 
    ylab('Signal') +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc)) +
    theme(legend.position = "none")

```




Make a subset which should be the linear range, plot again

```{r fig.height = 20, fig.width = 10}
SLOPE_FIRST_CYCLE <- 10
SLOPE_LAST_CYCLE <- 30


all.df.slope.subset <- all.df[all.df$cycle >= SLOPE_FIRST_CYCLE & all.df$cycle <= SLOPE_LAST_CYCLE, ]

ggplot(all.df.slope.subset, aes(x=cycle, y=rox_value, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Selected linear range') +
    xlab('Cycle') + 
    ylab('Signal') +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc)) +
    theme(legend.position = "none")

```


Subtract fluorescence from the no enzyme control

```{r fig.height = 20, fig.width = 10}
control.df <- all.df.slope.subset[all.df.slope.subset$final_enzyme_conc==0, c('assay_enzyme', 'sample_substrate', 'final_substrate_conc', 'cycle', 'repeat_num', 'rox_value')]

# rename the rox value
colnames(control.df)[length(colnames(control.df))] <- 'rox_value_control'

# do a left join to add on the controls to the samples
all.df.slope.subset.combined <- merge(all.df.slope.subset, control.df, by=(c('assay_enzyme', 'sample_substrate', 'final_substrate_conc', 'cycle', 'repeat_num')), all.x=TRUE)

# now calculate the normalized
all.df.slope.subset.combined$rox_value_norm <- all.df.slope.subset.combined$rox_value - all.df.slope.subset.combined$rox_value_control


ggplot(all.df.slope.subset.combined, aes(x=cycle, y=rox_value_norm, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Selected linear range, normalized data') +
    xlab('Cycle') + 
    ylab('Signal') +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc)) + 
    theme(legend.position = "none")

```



** Here I need to filter out the samples where all values fall below 0 **
```{r}

```



Fit a linear regression for all data between the first and last points, identify the one with the greatest R2 (or median or mean or something). This should be the one which is most linear

```{r}
# remove the control samples, they are not needed anymore as the data is now normalized
all.df.slope.subset.combined <- all.df.slope.subset.combined[!all.df.slope.subset.combined$final_enzyme_conc==0, ]


# first I need to bundle all the measurements from one well
all.df.slope.subset.combined.grouped <- all.df.slope.subset.combined %>%
    group_by(final_well, assay_well, sample_well, final_enzyme_conc, final_substrate_conc, repeat_num, assay_enzyme, sample_substrate) %>%
    summarize(cycle=list(cycle), 
              fam_value=list(fam_value), 
              rox_value=list(rox_value))

 
# now calculate the linear regression of ROX for each of the curves
slopes.df <- all.df.slope.subset.combined.grouped %>%
    group_by(final_well, assay_well, sample_well, final_enzyme_conc, final_substrate_conc, repeat_num, assay_enzyme, sample_substrate, final_substrate_conc) %>%
    mutate(rox_slope = coef( lm(unlist(rox_value)~unlist(cycle)) )[2],
              rox_r2 = summary( lm(unlist(rox_value)~unlist(cycle)) )$r.squared,
              fam_slope = coef( lm(unlist(fam_value)~unlist(cycle)) )[2],
              fam_r2 = summary( lm(unlist(fam_value)~unlist(cycle)) )$r.squared)

head(slopes.df)
```




A look at the slope data as a heatmap.

```{r}

### Plotting slope ###

# make a subset of the data frames for plotting
slope_rox_data.df <- slopes.df[, which(names(slopes.df) %in% c('assay_well', 'sample_well', 'rox_slope'))]

# make wide format
slope_rox_data_wide.df <- dcast(slope_rox_data.df, assay_well ~ sample_well, value.var='rox_slope')

# use first column as row names
rox.mat_data <- data.matrix(slope_rox_data_wide.df[, 2:length(names(slope_rox_data_wide.df))]) 
rnames <- slope_rox_data_wide.df[,'assay_well']   # assign labels in the "File" column to "rnames"
rownames(rox.mat_data) <- rnames 


# Now make the heatmap
myImagePlot(rox.mat_data, max=max(rox.mat_data), min=min(rox.mat_data), title=('ROX channel slope (linear model)'))




### Plotting R2 of linear model ###

# make a subset of the data frames for plotting 
df <- slopes.df[, which(names(slopes.df) %in% c('assay_well', 'sample_well', 'rox_r2'))]


# make wide format
df.wide <- dcast(df, assay_well ~ sample_well, value.var='rox_r2')

# use first column as row names
mat_data <- data.matrix(df.wide[, 2:length(names(df.wide))]) 
rnames <- df.wide[,'assay_well']   # assign labels in the "File" column to "rnames"
rownames(mat_data) <- rnames 

# Now make the heatmap
myImagePlot(mat_data, max=max(mat_data), min=min(mat_data), title=('ROX channel R2 of slope (linear model)'))


```


Remove all enzyme concentrations that have poor R2 values (below a set cutoff)

```{r}

# first I need to bundle all the measurements from each enzyme concentration
bundled.slopes.df <- slopes.df %>%
    group_by(assay_well, final_enzyme_conc, repeat_num, assay_enzyme, sample_substrate) %>%
    summarize(rox_slope=list(rox_slope), 
              rox_r2=list(rox_r2),
              final_substrate_conc=list(final_substrate_conc))


# now check, for each enzyme concentration, whether any of the R2 scores are below the cutoff
# a bad R2 score indicates a poor fit of the data to the linear model
bundled.slopes.nonegs.df <- bundled.slopes.df %>%
    group_by(assay_well, final_enzyme_conc, repeat_num, assay_enzyme, sample_substrate) %>%
    mutate(bad_r2 = any(unlist(rox_r2) < R2_CUTOFF))

# now actually filter the data
bundled.slopes.nonegs.df <- bundled.slopes.nonegs.df[bundled.slopes.nonegs.df$bad_r2 == FALSE, ]


head(bundled.slopes.nonegs.df)
```




Calculate spearman correlation between substrate concentration and the measured slope (the kinetic slope)

```{r}

bundled.slopes.nonegs.spearman.df <- bundled.slopes.nonegs.df %>%
        group_by(assay_well, final_enzyme_conc, repeat_num, assay_enzyme, sample_substrate) %>%
        mutate(slope_spearman_pval = cor.test(unlist(final_substrate_conc), unlist(rox_slope),  method='spearman')$p.value,
        slope_spearman_rho = cor.test(unlist(final_substrate_conc), unlist(rox_slope), method='spearman')[[4]])

head(bundled.slopes.nonegs.spearman.df)

barplot(bundled.slopes.nonegs.spearman.df$slope_spearman_rho, 
        names=interaction(bundled.slopes.nonegs.spearman.df$final_enzyme_conc, c('repeat'), bundled.slopes.nonegs.spearman.df$repeat_num),
        ylim=c(-1, 1),
        xlab='final_enzyme_conc',
        ylab='spearman rho',
        main='Which enzyme conc. is best?\n(corr. of substrate conc. and kinetic slope)')

```




For each repeat choose the enzyme concentration with maximal spearman rho

```{r}
bundled.slopes.nonegs.spearman.max.df <- bundled.slopes.nonegs.spearman.df %>%
    group_by(assay_enzyme, repeat_num, sample_substrate) %>%
    top_n(1, slope_spearman_rho) %>%
    slice(which.max(slope_spearman_rho))


head(bundled.slopes.nonegs.spearman.max.df)

barplot(bundled.slopes.nonegs.spearman.max.df$slope_spearman_rho, 
        names=interaction(bundled.slopes.nonegs.spearman.max.df$final_enzyme_conc, c('repeat'), bundled.slopes.nonegs.spearman.max.df$repeat_num),
        ylim=c(-1, 1),
        xlab='final_enzyme_conc',
        ylab='spearman rho',
        main='Selected enzyme concentrations')
```




Unbundle the dataset (turn the lists into column values again)

```{r}
bundled.slopes.nonegs.spearman.max.filtered.df <- bundled.slopes.nonegs.spearman.max.df[, -which(names(bundled.slopes.nonegs.spearman.max.df) %in% c('slope_spearman_rho', 'bad_r2', 'slope_spearman_pval'))]

bundled.slopes.nonegs.spearman.max.filtered.unbundled.df <- unnest(bundled.slopes.nonegs.spearman.max.filtered.df)

head(bundled.slopes.nonegs.spearman.max.filtered.unbundled.df)
```




Subtract the zero substrate slope from all others

```{r}

slope.of.control.df <- bundled.slopes.nonegs.spearman.max.filtered.unbundled.df[bundled.slopes.nonegs.spearman.max.filtered.unbundled.df$final_substrate_conc==0, c('assay_enzyme', 'sample_substrate', 'repeat_num', 'rox_slope')]

# rename the rox value
colnames(slope.of.control.df)[length(colnames(slope.of.control.df))] <- 'rox_slope_control'

# do a left join to add on the controls to the samples
final.data.df <- merge(bundled.slopes.nonegs.spearman.max.filtered.unbundled.df, slope.of.control.df, by=(c('assay_enzyme', 'sample_substrate', 'repeat_num')), all.x=TRUE)

# now calculate the normalized
final.data.df$rox_slope_norm <- final.data.df$rox_slope - final.data.df$rox_slope_control


ggplot(final.data.df, aes(x=final_substrate_conc, y=rox_slope_norm))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, plate %s, array %s, background-subtracted data', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate conentration (%s)', unique(sample.df$sample_substrate_unit))) + 
    ylab('Signal') +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))


```




Remove bad values (low concentrations where no activity is seen, whithin a certain tolerance)

```{r}

# first remove values close to zero
final.data.filtered.df <- final.data.df[final.data.df$rox_slope_norm > 0 + TOLERANCE_AT_ZERO * max(final.data.df$rox_slope_norm), ]


# determine the substrate concentration at the maximal slope is obtained
max.df <- final.data.filtered.df %>%
  select(assay_enzyme, sample_substrate, final_substrate_conc, repeat_num, final_enzyme_conc, rox_slope_norm) %>%
  group_by(assay_enzyme, sample_substrate, repeat_num) %>%
  slice(which.max(rox_slope_norm))

# drop the rox data and rename the substrate concentration column
max.df <- max.df[, -which(names(max.df) %in% c('rox_slope_norm'))]
names(max.df)[grep("final_substrate_conc", colnames(max.df))] <- 'final_substrate_conc_at_max'

# join the data frames
final.data.filtered.df <- merge(final.data.filtered.df, max.df, by=(c('assay_enzyme', 'sample_substrate', 'repeat_num', 'final_enzyme_conc')), all.x=TRUE)

# now remove values on the plateau if they are lower than the maximum rate, within some tolerance
final.data.filtered.both.df <- final.data.filtered.df %>%
  select(assay_enzyme, sample_substrate, final_substrate_conc, repeat_num, final_enzyme_conc, rox_slope_norm, final_substrate_conc_at_max) %>%
  group_by(assay_enzyme, sample_substrate, repeat_num) %>%
  filter(final_substrate_conc <= final_substrate_conc_at_max |
         final_substrate_conc > final_substrate_conc_at_max & rox_slope_norm >= max(rox_slope_norm) - TOLERANCE_AT_VMAX * max(rox_slope_norm))


# plot to see the data
ggplot(final.data.filtered.both.df, aes(x=final_substrate_conc, y=rox_slope_norm))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, plate %s, array %s, background-subtracted and filtered data', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate conentration (%s)', unique(sample.df$sample_substrate_unit))) + 
    ylab('Signal') +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))



```



Normalize data to the enzyme dilution

```{r}
final.data.filtered.both.df$rox_slope_norm_1x <- final.data.filtered.both.df$rox_slope_norm / final.data.filtered.both.df$final_enzyme_conc

head(final.data.filtered.both.df)
```



Plot slope vs substrate concentration, estimate Km and Vmax by fitting to the Michaelis-Menten equation

```{r}
in_data <- final.data.filtered.both.df


## Need to generalize this for many enzymes.... ##


# "velocity = Vmax times S divided by (Km plus S)", stored in MMcurve
MMcurve<-formula(v ~ Vmax*S/(Km+S))


# fit the equation to data using non-linear least squares
kinetic_data <- in_data[, which(names(in_data) %in% c('assay_enzyme', 'rox_slope_norm_1x', 'final_substrate_conc'))]

# rename to make it work with the formula
slope_index <- grep("rox_slope_norm_1x", colnames(kinetic_data))
names(kinetic_data)[slope_index] <- 'v'

substrate_index <- grep("final_substrate_conc", colnames(kinetic_data))
names(kinetic_data)[substrate_index] <- 'S'


# estimate Vmax
vmax_estimate <- max(kinetic_data$v)


# estimate km at concentration closest to half of Vmax
index_of_half_vmax <- which.min(abs(kinetic_data$v - max(kinetic_data$v)/2))
km_estimate = kinetic_data$S[index_of_half_vmax]
  

# perform non-linear least squares regression
bestfit <- nls(MMcurve, kinetic_data, start=list(Vmax=vmax_estimate, Km=km_estimate))


# generate new x-values that can be used to predict y using the model (for making the smooth curve)
xNew <- seq(0, max(in_data$final_substrate_conc), length.out = 100)
yNew <- predict(bestfit, newdata=list(S = xNew))

p1 <- ggplot(in_data, aes(x=final_substrate_conc, y=rox_slope_norm_1x))+
    theme_bw() +
    geom_point(cex=1.5, fill='red', color='red') + 
    geom_line(data=data.frame(xNew, yNew), aes(x=xNew, y=yNew)) + 
    ggtitle(sprintf('Km plot, plate %s, array %s', PLATE_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate (%s)', unique(sample.df$sample_substrate_unit))) + 
    ylab('Signal') +
    annotate("text", x = max(in_data$final_substrate_conc) - max(in_data$final_substrate_conc)*0.1, y = min(in_data$rox_slope_norm_1x) + max(in_data$rox_slope_norm_1x)*0.05, label = sprintf("Vmax = %s", formatC(coef(bestfit)['Vmax'], format = "e", digits = 2))) + 
    annotate("text", x = max(in_data$final_substrate_conc) - max(in_data$final_substrate_conc)*0.1, y = min(in_data$rox_slope_norm_1x) + max(in_data$rox_slope_norm_1x)*0.1, label = sprintf("Km = %s %s", round(coef(bestfit)['Km'], 2), unique(sample.df$sample_substrate_unit))) + 
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme))


p1

# save the figure
fname = paste(FINAL, THIS_PROJECT_FOLDER, sprintf('plate_%s_array_%s_Km_plot.pdf', PLATE_NUMBER, ARRAY_NUMBER), sep='/')
ggsave(filename=fname, plot=p1, width = 150*length(unique(sample.df$sample_substrate)), height = 100*length(unique(assay.df$assay_enzyme)), units=c('mm'))


# save the km and kcat

```


