---
output:
  html_document: default
  pdf_document: default
---

---
title: "Test of whether kinetic constants can be obtained with BioMark"
output: html_notebook
---

Copyright (C) 2020  Martin Engqvist Lab




## Experiment
.....

Experimental data collected on 160719 (DDMMYY) in Petter Wolls group at Karolinska using a BioMark instrument.

Assay conditions, oxidases:


20mM HEPES, pH 7.4

50uM Amplex Red

0,1U/ml HRP

0,1uM fluorescein

0,01mg/ml BSA

0,01% Triton-X-100 and Tween20


```{r setup}

#make sure the packages are installed
wants <- c('dplyr', 'reshape2', 'ggplot2', 'tidyverse', 'assertthat')
has   <- wants %in% rownames(installed.packages())
if(any(!has)) install.packages(wants[!has], repos="http://ftp.acc.umu.se/mirror/CRAN/")

#load all the packages
lapply(wants, require, character.only = TRUE)


#find the "code" directory, this should be one folder away from the project root
while (basename(getwd()) != 'code'){
  setwd("../")
  if (getwd() == '/'){
    quit(status = 1)
  }
}

#go one folder up and set as project root directory
setwd("../")
PROJ <- getwd()
setwd(paste0(PROJ, '/data/'))

#set folder paths relative to the project root
CURRENT <- getwd()
DATA <- getwd()
RAW_EXTERNAL <- paste0(DATA, '/raw_external/')
RAW_INTERNAL <- paste0(DATA, '/raw_internal/')
INTERMEDIATE = paste0(DATA, '/intermediate/')
FINAL = paste0(DATA, '/final/')

RESULTS <- paste0(PROJ, '/results/')
FIGURES <- paste0(RESULTS, 'figures/')
PICTURES <- paste0(RESULTS, 'pictures/')

```


## Define variables specific for this chip and array
```{r}
THIS_PROJECT_FOLDER <- 'flex_six_km_test'
CHIP_NUMBER <- 4
ARRAY_NUMBER <- 1
DATA_FOLDER <- sprintf('chip_%s_Array_%s', CHIP_NUMBER, ARRAY_NUMBER)
FAM_DATA <- '1530024197_chip_4_array_1_fam.tsv'
ROX_DATA <- '1530024197_chip_4_array_1_rox.tsv'

MINUTES_PER_CYCLE <- 0.5 # how many minutes elapse between each cycle

R2_CUTOFF <- 0.8
CORR_CUTOFF <- 0.8
TOLERANCE_AT_ZERO <- 0.05
TOLERANCE_AT_VMAX <- 0.20

SLOPE_FIRST_TIMEPOINT <- 0
SLOPE_LAST_TIMEPOINT <- 15


# load helper functions
source(paste0(PROJ, '/code/', 'flex_six_helper_functions.R'))
print(paste0('helper script version: ', VERSION))
```


Setup plotting preferences

```{r}
font_size = 7
plt_height = 65
plt_width = 89
line_size = 0.5
line_width <- 0.5
error_size = 0.25
pd = position_dodge(0.9)

gen_col = '#141A46'
nat_col = '#EC8B5E'
neg_col = '#868686'



ggplot_theme <- theme(text = element_text( size=font_size), 
                      axis.text.x = element_text(angle = 0, hjust = 1, size=font_size),
                      axis.text.y = element_text(size=font_size),
                      axis.line = element_line(color = "black", size=line_size),
                      axis.ticks.x = element_line(color = 'black', size=line_size),
                      axis.ticks.y = element_line(color = 'black', size=line_size),
                      #legend.text = element_text( size=font_size),
                      legend.position = c(0, 1),
                      legend.justification = c(0, 1),
                      legend.background = element_blank(),
                      legend.direction = 'horizontal',
                      legend.text=element_text(size=font_size),
                      plot.title = element_text( size=font_size),
                      panel.background = element_blank(),
                      panel.border = element_rect(color = 'black', size=line_size),
                      #panel.margin = 
                      panel.grid.major = element_blank(), 
                      panel.grid.minor = element_blank(),
                      panel.grid.major.x = element_blank(), 
                      panel.grid.major.y = element_blank(), 
                      panel.grid.minor.x = element_blank(), 
                      panel.grid.minor.y = element_blank() 
)
```



Setup

Let's input the experimental paramenters for sample and assay sides

```{r}
assay.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, sprintf('chip%s_array%s_assay.tsv', CHIP_NUMBER, ARRAY_NUMBER), sep='/'), sep='\t', header=TRUE, na.strings="")

sample.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, sprintf('chip%s_array%s_sample.tsv', CHIP_NUMBER, ARRAY_NUMBER), sep='/'), sep='\t', header=TRUE, na.strings="")


print(assay.df)
print(sample.df)
```



"Mix" the wells from input assay and sample data to do in silico what happens in the chip

```{r}
mixed.df <- mix_wells(assay.df, sample.df)

head(mixed.df)
tail(mixed.df)
```



Load data

```{r}
# read the experimental data obtained from the BioMark
rox.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, ROX_DATA, sep='/'), sep='\t', header=TRUE, na.strings="")
fam.df <- read.table(paste(INTERMEDIATE, THIS_PROJECT_FOLDER, FAM_DATA, sep='/'), sep='\t', header=TRUE, na.strings="")

# combine data 
all.df <- combine_data(mixed.df, rox.df, fam.df)

# Add in the time at which each read took place
all.df$minutes <- all.df$cycle * MINUTES_PER_CYCLE

head(all.df)
```



Take a look at the frame and remove NA if required

```{r}
summary(all.df)

# replace all na values with zero to allow for analysis despite missing values
all.df[is.na(all.df)] <- 0

```



Let's export this to a flatfile for future use.

```{r}
write.table(all.df, file=paste(FINAL, THIS_PROJECT_FOLDER, sprintf('%s_flex_six_chip_%s_array_%s_raw_data.tsv', gsub(' ', '_', unique(all.df$assay_enzyme)), CHIP_NUMBER, ARRAY_NUMBER), sep='/'), sep='\t', row.names=FALSE)
```



## Plotting results
Time to plot!

Plot rox and fam values of first and last cycle

```{r}

plot_raw_data_heatmaps(all.df, min=0, max=65000)

```



First we plot the raw kinetics observed in all the reaction chambers

```{r}
ggplot(all.df, aes(x=minutes, y=rox_value, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Raw data measured over time, chip %s, array %s', CHIP_NUMBER, ARRAY_NUMBER)) +
    xlab('Time (minutes)') + 
    ylab('Fluorescence') +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```



Subtract fluorescence from the no enzyme control

```{r}
# remove fluorescence from the no enzyme control
all.df <- subtract_no_enz_fluorescence(all.df)

head(all.df)

in_data <- all.df
ggplot(in_data, aes(x=minutes, y=rox_value_norm, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, chip %s, array %s, normalized data', CHIP_NUMBER, ARRAY_NUMBER)) +
    xlab('Time (minutes)') + 
    ylab(sprintf('Resorufin concentration (%s)', unique(in_data$resorufin_value_norm_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```
These look quite horrible. That is because the fluorescecne of the no enzyme control increases with time combined with the fact that some samples maxed out the instrument in terms of intensity, so when normalizing one gets this decreasing slope. This should be fine, however, as we will only use initial rates for the "best" enzyme concentration, so the sagging part of the lines will not be analyzed.



Convert fluorescence to concentration of product using standard curve

```{r}
concentration.df <- fluorescence_to_concentration(all.df, chip_type='flex_six')

head(concentration.df)

in_data <- concentration.df
ggplot(in_data, aes(x=minutes, y=resorufin_value_norm, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Raw data measured over time, chip %s, array %s', CHIP_NUMBER, ARRAY_NUMBER)) +
    xlab('Time (minutes)') + 
    ylab(sprintf('Resorufin concentration (%s)', unique(in_data$resorufin_value_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```



Remove substrate concentrations where no activity is seen  ## New block ##

```{r}
# Remove bad enzyme-substrate combinations (low concentrations where no activity is seen)
concentration.df <- trim_data_low_range(concentration.df, tolerance_at_zero=TOLERANCE_AT_ZERO)

# look at the data
head(concentration.df)


# plot it
in_data <- concentration.df
ggplot(in_data, aes(x=minutes, y=resorufin_value_norm, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle('Raw data measured over time') +
    xlab('Time (minutes)') + 
    ylab(sprintf('Substrate consumed (%s)', unique(in_data$final_substrate_norm_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```



Make a subset which should be the linear range, plot again

```{r}
all.df.slope.subset <- concentration.df[concentration.df$minutes >= SLOPE_FIRST_TIMEPOINT & concentration.df$minutes <= SLOPE_LAST_TIMEPOINT, ]

in_data <- all.df.slope.subset
ggplot(in_data, aes(x=minutes, y=resorufin_value_norm, color=factor(final_substrate_conc)))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, chip %s, array %s', CHIP_NUMBER, ARRAY_NUMBER)) +
    xlab('Time (minutes)') + 
    ylab(sprintf('Resorufin concentration (%s)', unique(in_data$resorufin_value_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))

```



Fit a linear regression for all data between the first and last points to get the slope.

```{r}
# Fit a linear regression for all data between the first and last points to get the slope.
slopes.df <- compute_slopes(all.df.slope.subset)

head(slopes.df)
tail(slopes.df)
```



A look at the slope data as a heatmap and assess the fit of the linear model.

```{r}
plot_slope_data_heatmaps(slopes.df)
```



Remove all enzyme concentrations that have poor R2 values (below a set cutoff)

```{r}

bundled.slopes.nonegs.df <- filter_r2(slopes.df, r2_cutoff=R2_CUTOFF, corr_cutoff=CORR_CUTOFF)


head(bundled.slopes.nonegs.df)
```



Calculate spearman correlation between substrate concentration and the measured slope (the kinetic slope), the 
idea being that we expect higher substrate concentraions to lead to a steeper slopes.
```{r}

bundled.slopes.nonegs.spearman.df <- calculate_slope_and_substrate_correlation(bundled.slopes.nonegs.df)

# look at the data
head(bundled.slopes.nonegs.spearman.df)

# plot it
barplot(bundled.slopes.nonegs.spearman.df$slope_spearman_rho, 
        names=interaction(bundled.slopes.nonegs.spearman.df$final_enzyme_conc, c('repeat'), bundled.slopes.nonegs.spearman.df$repeat_num),
        ylim=c(0, 1),
        xlab='final_enzyme_conc',
        ylab='spearman rho',
        main='Which enzyme conc. is best?\n(corr. of substrate conc. and kinetic slope)',
        las=3)

```



For each repeat choose the enzyme concentration with maximal spearman rho

```{r}

bundled.slopes.nonegs.spearman.max.df <- select_best_samples(bundled.slopes.nonegs.spearman.df)

# look at the data
head(bundled.slopes.nonegs.spearman.max.df)

# plot it
barplot(bundled.slopes.nonegs.spearman.max.df$slope_spearman_rho, 
        names=interaction(bundled.slopes.nonegs.spearman.max.df$final_enzyme_conc, c('repeat'), bundled.slopes.nonegs.spearman.max.df$repeat_num),
        ylim=c(0, 1),
        xlab='final_enzyme_conc',
        ylab='spearman rho',
        main='Selected enzyme concentrations',
        las=3)
```



Unbundle the dataset (turn the lists into column values again)

```{r}
bundled.slopes.nonegs.spearman.max.filtered.unbundled.df <- unbundle_data(bundled.slopes.nonegs.spearman.max.df)

head(bundled.slopes.nonegs.spearman.max.filtered.unbundled.df)
```



Need to convert from uM per minute to micro moles per minute

```{r}
final.data.moles.df <- concentration_to_moles(bundled.slopes.nonegs.spearman.max.filtered.unbundled.df, chip_type='flex_six')

head(final.data.moles.df)
```



Normalize to per mg protein

```{r}
final.data.moles.protein.df <- normalize_to_protein(final.data.moles.df, chip_type='flex_six')

head(final.data.moles.protein.df)
```



Make a plot of the new values

```{r}
# plot it
in_data <- final.data.moles.protein.df
ggplot(in_data, aes(x=final_substrate_conc, y=activity_moles_protein))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Activity, chip %s, array %s', CHIP_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate conentration (%s)', unique(sample.df$sample_substrate_unit))) + 
    ylab(sprintf('Activity (%s)', unique(in_data$activity_moles_protein_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))
```



Trim off bad values at very high substrate concentrations  ## Block changed ##

```{r}

# Remove bad values (low concentrations where no activity is seen, and high values where activity drops below max)
final.data.filtered.both.df <- trim_data_top_range(final.data.moles.protein.df, tolerance_at_max=TOLERANCE_AT_VMAX)

# look at the data
head(final.data.filtered.both.df)


# plot it
in_data <- final.data.filtered.both.df
ggplot(in_data, aes(x=final_substrate_conc, y=activity_moles_protein))+
    theme_bw() +
    geom_point(cex=0.5) + 
    ggtitle(sprintf('Selected linear range, chip %s, array %s, filtered data', CHIP_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('Substrate conentration (%s)', unique(sample.df$sample_substrate_unit))) + 
    ylab(sprintf('Activity (%s)', unique(in_data$activity_moles_protein_unit))) +
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme, final_enzyme_conc))



```



Export for future use

```{r}
write.table(final.data.filtered.both.df, file=paste(FINAL, THIS_PROJECT_FOLDER, sprintf('%s_flex_six_chip_%s_array_%s_final_data.tsv', gsub(' ', '_', unique(all.df$assay_enzyme)), CHIP_NUMBER, ARRAY_NUMBER), sep='/'), sep='\t', row.names=FALSE)
```



Plot slope vs substrate concentration, estimate Km and Vmax by fitting to the Michaelis-Menten equation

```{r}

kinetic_data <- calculate_kinetics(final.data.filtered.both.df)
mm_fit.df <- data.frame(xFit=kinetic_data$xFit, yFit=kinetic_data$yFit)
km <- kinetic_data[[3]]
vmax <- kinetic_data[[4]]


in_data <- final.data.filtered.both.df

p1 <- ggplot(in_data, aes(x=final_substrate_conc, y=activity_moles_protein))+
    theme_bw() +
    ggplot_theme +
    geom_point(cex=1.5, fill='red', color='red') + 
    geom_line(data=mm_fit.df, aes(x=xFit, y=yFit)) + 
    ggtitle(sprintf('Km plot, chip %s, array %s', CHIP_NUMBER, ARRAY_NUMBER)) +
    xlab(sprintf('%s (%s)', unique(in_data$sample_substrate), unique(in_data$final_substrate_unit))) + 
    ylab( expression(paste('Activity (', mu, 'mol mg'^{-1}, 'min'^{-1}, ')') ) ) +

    annotate("text", 
            x = min(in_data$final_substrate_conc) + max(in_data$final_substrate_conc)*0.2, 
            y = min(in_data$activity_moles_protein) + max(in_data$activity_moles_protein)*0.05, 
            label = as.expression(bquote(italic(V)[max] ~ '=' ~ .(formatC(vmax, format='g', digits=2)) ~ mu * 'mol mg'^{-1} ~ 'min'^{-1})), 
            size=2,
            hjust=0,
            vjust=0) +
  
    annotate("text", 
             x = min(in_data$final_substrate_conc) + max(in_data$final_substrate_conc)*0.2, 
             y = min(in_data$activity_moles_protein) + max(in_data$activity_moles_protein)*0.15, 
             label = as.expression(bquote(italic(K)[M] ~ '=' ~ .(formatC(km, format='g', digits=2)) ~ 'mM')), 
             size=2,
             hjust=0,
             vjust=0) + 
  
    facet_grid(cols=vars(sample_substrate), rows=vars(assay_enzyme))


p1

# save the figure
fname = paste(FIGURES, sprintf('%s_flex_six_chip_%s_array_%s_Km_plot.pdf', gsub(' ', '_', unique(all.df$assay_enzyme)), CHIP_NUMBER, ARRAY_NUMBER), sep='/')
ggsave(filename=fname, plot=p1, width = 65*length(unique(sample.df$sample_substrate)), height = 50*length(unique(assay.df$assay_enzyme)), units=c('mm'))


# save the km and kcat

```